---
title: Gendered words over time in the Times Data Archive
---


We can start by loading the libraries we will be using for data analysis. 

```{r loadLibraries}
suppressPackageStartupMessages(library(yaml))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(pander))
suppressPackageStartupMessages(library(car))
```


Now we need to load in our data file and change the structure from YAML
to something that we can more easily manipulate.

```{r loadData}
data = yaml.load_file("result.yml")
counts = lapply(names(data), function(year){
  lapply(data[[year]], function(pair) {
    data.frame(year=as.numeric(year), word=pair[[1]], count=as.numeric(pair[[2]]))
  }) %>% rbindlist
}) %>% rbindlist
```

We should merge the multiple spelling words. These are:

   * Househusband, house husband, house-husband
   * Housewife, house wife, house-wife

These may be the only words which contain "-" or " ", which we should check.

```{r checkForHyphenOrSpace}
counts %>% 
  ungroup %>%
  dplyr::select(word) %>% 
  unique %>%
  filter(grepl("[ -]", word)) %>% 
  as.vector() %>% 
  pander
```

These are the only words that contain "-", but "he said" and "she said" contain " ". So we should be a bit more careful and use `recode`. However, after
this we will need to recompute the counts.

```{r mergeMultiWords}
counts %<>%
  mutate(word = recode(word, "'house husband'='househusband';'house-husband'='househusband';'house wife'='housewife';'house-wife'='housewife'")) %>%
  group_by(year, word) %>%
  summarize(count = sum(count))
```
 

```{r simplePlot,fig.height=50,fig.width=10}
ggplot(counts, aes(x=year,y=count +1)) + 
  geom_point() +
  geom_line() +
  scale_y_log10() +
  facet_wrap(~word, ncol=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

There are a few things that I have noticed that might be interesting:

   * himself/herself/itself etc only seem to appear in the archive at the start of the 1800's
   * The length of the times archive per year / issue / article is not constant. This will hide aspects of the trend. 
   
We can start by trying to address the issue of word published per year. Fortunately, we have this data available as well.

```{r getYearWordCounts}
words_per_year = fread('words_per_year.csv', header=T)
```

And we can see there is a large difference in how many words are published year to year.

```{r showWordsIncreasingAnnually}
ggplot(words_per_year,aes(year,num_words)) +
  geom_point() + 
  geom_line()
```

We can now add this data into our table of occurances.

```{r joinCountsandTotalCounts}
combined = left_join(counts,words_per_year,by="year")
assertthat::are_equal(nrow(counts),nrow(combined))
```

This enables to us to plot how much of the total annual writing a single word makes up over time.
This normalises for the effect of the changing size of the corpus from year to year.

Since the words all appear with very differing frequencies, we will make the plots each have 
independant x axes, so we can see the trend for each word individually.

```{r normalisedWordCountPlot, fig.height=50, fig.width=10}
ggplot(combined, aes(year, 10000 * count/num_words)) +
  geom_point() + 
  geom_line() +
  xlab("Year") + 
  ylab("Per 10k of annual words") +
  facet_wrap(~word, ncol=3,scales = "free_y")
```
 